changes since submission: 
1. changed 56,000 to 55,000 
2. typographical errors corrected
3. added reference to Bailer Jones 2013
4. updated figures 8 and 9 

55,000 stars as there were some duplicates in the list -there are 57454 unique id's in the DR10 allStar files with parameters but not all of these spectra are available online - downloaded the 55,000 available in the fields to run through. 

1) Line 670:  removed after submitted: `` Labels are provided for stars with both \aspcap-corrected labels''. Also fixed some small errors in the table formatting: Two columns were around the wrong way, for the TARG2 and ASPCAPFLAG and also there was a "K" in units under the log g. - see (3) 
 
 2 Title: change "for" to "to"
 
 3) Line 676: {2MASS} &  & K &  dex  & K & dex & dex & dex & kms$^{-1}$ & &  \tiny{WARN} & \tiny{ FLAG}  \\    
 
 4) Line 122 full stop after references. 
 
 To do: 
 
 Need to have a look at the x-over stars that are observed in commissioning mode and also deal with this in star count (1700 stars). 
 
 Also, was thinking - maybe it is better to not do the flag cut on STAR BAD flag as this is a catch all flag coming from ASPCAP- maybe a more 'appropriate' flags cut is the following:
 1. 2**25  in aspcapflag with is colour spectroscopic far away from colour photometric warning
 2. 2**26 in aspcapflag which is rotation bad (or 2**10 which is rotation warning)
 3. 2**9 in target2 flag which is telluric
 4. vscatter < 10 km/s
 which gives 38,000 stars in these figures. 
 
 This has more stars at high log g in the aspcap-corrected results but the isochrone-corrected results looks the same and there are 38,000 stars compared to 35,000 in the current figures. 
 
 5. stupidly had error in mkn_TCA_v1_edit in that a star in M2, M3 and M13 was on main sequence rather than being eliminated from this figure. removed these and remade figure. also reran training through with these fixed. 
 
5. CHi2 statistic is not done properly- not true i think it is : overrule this I divide by this number but include this number - have divided by number of pixels when need to divide by number of useable pixels = 7200 - labels fit so chi2 will increase  - can do this retrospectively
though, just multiply all values by 1.15 - do this for every value  - note high chi2 is not necessarily a bad thing in fact this may be expected if non label dependent regions significantly diverge
due to individual elements  . No chi2 should be good shouldn't need to change this. 

these changes went into the version posted on astroph. 

6. Anh p 6 extra )
7. only cross out p. 7 
20. [$\alpha$/Fe] 
21: penal - panel 
22.in -> on
22. C: 
23: D -> DR10
23: Object -> objects
23: cross out are
23: \logg to \logg\ 
31: Pleiades from Hyades
32: effect -> affect
33. had indicates -> had indicated
17: in in \logg\
22. removed in.
25. missing 4800 K temp. 
19 removed covariant. full stop

21. change Anna's initials: Anna Y. Q. Ho 

To do: figure 8 - change this so that it is 5600 K stopping
morgan had some suggestions regarding plots may wish to implement these 

March 15 2014
-- had to change abstract so it is shorter. \\
