% This file is part of the Apogee project.
% Copyright 2014 the authors.

\documentclass[12pt, preprint]{aastex}
\input{vc}

\newcommand{\sectionname}{Section}
\newcommand{\project}[1]{\textsl{#1}}

\begin{document}

\title{Data-driven chemical tags for stars}
\author{
  some combination of DWH, MKN, HWR, and others
}
\date{DRAFT / \gitdate\ / \githash\ / NOT FOR DISTRIBUTION}

\begin{abstract}
Stars with common origins can be found if they can be tagged with
sufficiently informative tags (chemical and other properties) that are
invariant with time from birth to the present day.
We perform an uninterpreted tagging of stars without the direct use of
stellar models or model spectra from libraries.
The principle is to construct temperature- and gravity-insenstive
proximity measures in spectral space, and perform unsupervised
clustering.
Work on real and simulated spectra are promising.
\end{abstract}

\section{introduction}

The idea behind chemical tagging is that every stellar birth
environment (every molecular cloud, say) is expected to have slightly
different chemical abundances in detail from every other birth
environment, such that stars with common origins but different
present-day phase-space locations can in principle be identified by
their surface chemical abundances (and possibly also ages).
This potentially revolutionary scientific program has not yet
delivered results, in part because it seems to require tens or
hundreds of thousands of high resolution, high signal-to-noise stellar
spectra, and in part because it requires stellar models good enough to
make consistent measurements of abundances across a wide range of
spectral types (temperatures and surface gravities).

There are a range of new projects obtaining the required spectra,
including \project{APOGEE} (cite) and \project{HERMES} (cite).
There are many efforts to analyze and improve stellar models (cite).
However, it is possible at the present day---with spectra in
hand and without trustworthy models at all---to assess the feasiblity
of chemical tagging projects.
The idea is the following:
If sufficient information resides within high resolution and high
signal-to-noise spectra to deliver informative tags, then that
information will be visible in the spectra \emph{themselves}.
To see that the spectra are informative does not (necessarily) require
accurate models of those spectra.
Models might be required to \emph{interpret} the information, but they
are not required to \emph{identify its existence}.
And of course the contrapositive is also true:
Since abundance measurements are made using the spectral data, if the
information to perform precise chemical tagging is \emph{not} present
in the spectra themselves, it cannot be present in any abundance
measurements derived therefrom.

In what follows, we use (simple, straightforward) machine-learning
methods that have been designed to find structure in large, high
dimensional data sets to look for structure in stellar spectra.
We design these methods to be maximally sensitive to stellar
properties \emph{other} than temperature and surface gravity, since
these properties have enormous range within any individual stellar
population.
The hope is that structure seen in the ``tangent space'' (as it were)
to these properties will be some transformation of structure in the
chemical properties of the stars (and ages).

...Related prior work includes PCA, HMF, XD, SVM.  These all build
data-driven models of non-trivial data sets, and all could be used
here.  We are going to try to stick with methods that involve building
probabilistic generative models, or approximations thereto...

\section{generalities}

Data-driven models can be implemented for problems like this in a
number of ways.  In this \sectionname, we describe three formulations
of the problems---as a set of classifications, as a regression, and as
density estimation in data space---and remark on technical approaches
for each formulation.

\paragraph{classification formulation:}
...Classification description of the problem.  We have stars from two
clusters; we want to find the directions or subspace of the data space
in which the clusters are \emph{predictably} best classified.  This
method is supervised in that it requires cluster identities up-front
(for at least some stars).  The harder classification problem is to
take stars from one cluster and then stars from \emph{all other
  sources}.  In what subspace are the stars from the cluster separated
from everything else?  As for methods: Compare SVM with a (modified,
heteroskedastic) Factor Analyzer; why is some modification of the FA
so much better for this problem?

\paragraph{regression formulation:}
...Regression description of the problem.  We have stars in a range of
ages or alpha-abundances, and we either start with known ages or alpha
enhancements (the supervised version of the problem) or else we start
with an emission-line or absorption-line profile that indicates age or
alpha enhancement (the unsupervised version of the problem).  In the
supervised version, we fit every star as a base spectrum plus age
indicator components.  In the unsupervised version, we fit every star
as a base spectrum plus known profile.  The base spectrum, in each
case, is a data-driven model, or the best you can do with all the data
you have ever seen.  HMF is good for this.  But if you want to
marginalize out stellar base spectrum, then you need to build priors.
Hierarchical is the best way to do that.

\paragraph{density estimation formulation:}
...Density estimation description of the problem.  We have a large
number of stars and we try to describe the distribution in spectral
space with a mixture of lines or hypersurfaces.  Show that the stars
lie on a finite number (hundreds or thousands) of well-separated
hypersurfaces and don't fill the bulk.  That's hard, but might be
possible if you use existing models as a guide or initialization.

\section{data and objectives}

\section{results and discussion}

\acknowledgements It is a pleasure to thank
  Ross Fadely,
  Rob Fergus,
  Dan Foreman-Mackey
for valuable discussions.

\end{document}
